Left-Linear Rewriting in Adhesive Categories

### 2024-07-02 concur/submitted/fusione.tex

  Two derivations $\der{D}, \der{E}:G\Mapsto H$ are \emph{switch
    equivalent}, written $\der{D}\equiv^{sh} \der{E}$, if there is a
  switching sequence $\{\der{D}_{k}\}_{k=0}^m$ such that
  $\der{D}\equiv_a \der{D}_0 \shift{\nu_1} \der{D}_1 \shift{\nu_2}
  \ldots \shift{\nu_m} \der{D}_m \equiv _a \der{E}$.
\end{definition}

@@

In the DPO approach to rewriting, the notion of sequential
independence, identifying consecutive rewriting steps which are
independent and thus can be performed in reverse order is typically
closely related to that of parallel
independence~\cite{ehrig2014adhesive}. Parallel independence
characterises the independence of two rewriting steps starting from
the same object, and guarantees that the two steps can be applied in
any order producing, up to isomorphism, the same resulting
object. Parallel independent steps, when applied in sequence, lead to
sequential independent steps and, conversely, from two sequential
independent steps, as a byproduct of the switch construction, one can
get two coinitial parallel independent steps.  Under mild additional
conditions, one can have a notion of parallel rule in a way that
parallel independent steps can be also applied in parallel.
%
Even if this is not discussed explicitly in the paper, our theory can be easily accomodated to ecompass a notion of parallel independence and parallel rule application, enjoying the properties outlined above. To ensure this, however, one must require that the underlying category $\X$ has binary coproducts and that the class $\mathcal{M}$ is closed under them.

>>>>

  Two derivations $\der{D}, \der{E}:G\Mapsto H$ are \emph{switch
    equivalent}, written $\der{D}\equiv^{sh} \der{E}$, if there is a
  switching sequence $\{\der{D}_{k}\}_{k=0}^m$ such that
  $\der{D}\equiv_a \der{D}_0 \shift{\nu_1} \der{D}_1 \shift{\nu_2}
  \ldots \shift{\nu_m} \der{D}_m \equiv _a \der{E}$. 
  To point out a chosen permutation of rewriting steps, we will also write $\der{D}\equiv^{sh}_{\sigma} \der{E}$, where $\sigma$ is the composition of the transposition appearing in a chosen switching sequence. 
\end{definition}

@@

In the DPO approach to rewriting, the notion of sequential
independence, identifying consecutive rewriting steps which %are independent and thus
 can be performed in reverse order is typically
closely linked to that of \emph{parallel
independence}~\cite{ehrig2014adhesive}. Parallel independence
%characterises the independence of 
relates two rewriting steps starting from
the same object, %and guarantees that the two steps 
which can be applied in
any order producing, up to isomorphism, the same resulting
object. Parallel independent steps, when applied in sequence, lead to
sequential independent steps and, conversely, from two sequential
independent steps, as a byproduct of the switch construction, one can
get two %coinitial 
parallel independent steps.

Under mild additional
conditions, one can have a notion of \emph{parallel rule} in a way that two
parallel independent steps can be also %applied both 
combined in a single rewriting step. %in parallel.
%
Even if this is not discussed explicitly in the paper, our theory can be easily accomodated to ecompass a notion of parallel independence and parallel rule application, enjoying the properties outlined above. To ensure this, however, one must require that the underlying category $\X$ has binary coproducts and that the class $\mathcal{M}$ is closed under them.
